Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2021-03-15 Manuscript/HI_Model_ms.Rmd
updated manuscript files

### *4.3 Limitations of this study* 
There is ample evidence that increasing atmospheric CO~2~ concentrations can stimulate NPP [@norby_forest_2005; @zhu_greening_2016], but the magnitude and persistence of this effect across climatic and nutrient availability gradients remains highly uncertain, especially over long time spans [@franks_sensitivity_2013; @walker_integrating_2020]. Our results demonstrate that long-term projections of ecosystem carbon balance are highly sensitive to uncertainty in the strength of a CFE. With no CFE, Hawaiian ecosystems became a net source of CO~2~ to the atmosphere beginning in the latter half of the 21st century under the high land use and high radiative forcing scenario. However, a CFE equivalent to a 5% increase in NPP for every 100 ppm increase in atmospheric CO~2~ applied to the same scenario resulted in Hawaiian ecosystems collectively remaining a net carbon sink throughout the entire simulation period. A CFE of 15% applied to the high land use and high radiative forcing scenario resulted in a nearly 5-fold increase in mean annual NBP averaged across all years and Monte Carlo realizations.

>>>>

### *4.3 Limitations of this study* 
There is ample evidence that increasing atmospheric CO~2~ concentrations can stimulate NPP [@norby_forest_2005; @zhu_greening_2016], but the magnitude and persistence of this effect remains highly uncertain, especially over long time spans [@franks_sensitivity_2013; @walker_integrating_2020]. Our results demonstrate that long-term projections of ecosystem carbon balance are highly sensitive to uncertainty in CFE strength. With no CFE, Hawaiian ecosystems transitioned into a net source of CO~2~ to the atmosphere beginning in the latter half of the 21st century under the high land use and high radiative forcing scenario. However, a CFE equivalent to a 5% increase in NPP for every 100 ppm increase in atmospheric CO~2~ applied to the same scenario resulted in Hawaiian ecosystems remaining a net carbon sink throughout the entire simulation period. A 15% CFE applied to the high land use and high radiative forcing scenario resulted in a nearly 5-fold increase in mean annual NBP averaged across all years and Monte Carlo realizations. Despite this demonstrated sensitivity to a CFE, the combined effect of several potentially attenuating factors makes it difficult to determine the appropriate magnitude of a CFE with any degree of confidence [@walker_integrating_2020]. Nitrogen and phosphorus limitation can reduce or eliminate a CFE [@reich_nitrogen_2006; @norby_co2_2010; @he_nitrogen_2017; @terrer_nitrogen_2019], as can water limitation and heat stress [@obermeier_reduced_2017; @birami_hot_2020]. Forest age may also be a factor, with young aggrading forests showing a strong positive growth response to CO~2~ fertilization [@walker_decadal_2019], while old-growth forests show little to no response [@jiang_fate_2020; @yang_low_2020]. This evidence indicates that a CFE may be highly variable across space and time, suggesting that it may be unrealistic to apply a single CFE rate value across an entire region over several decades. Until mechanistic understanding is improved, the most conservative approach when projecting future ecosystem balance in the context of climate mitigation planning may be to assume no CFE, with the knowledge that any realized CFE will only enhance ecosystem carbon sequestration.
